### 2017-04-29 paper/jfp-submission/evaluation.tex
significance

@@

\subparagraph{Internal.}
%
We assigned students randomly to two groups. The first group was given
\ocaml's errors for @append@ and \hbox{@digitsOfInt@,} and \toolname's trace
for @sumList@ and @wwhile@. The second group was given the opposite
assignment of errors and traces. This assignment ensured that (1) each
student was given \ocaml and \toolname problems, and (2) each student
was given an ``easy'' and ``hard'' problem for both \ocaml and
\toolname. Students without sufficient knowledge of \ocaml could affect
the results, as could the time-constrained nature of an exam. For these
reasons we excluded any answers left blank from our analysis.

@@

\subparagraph{Conclusion.}
%
We collected exams from 60 students, though due to the nature of the
study not every student completed every problem.
%
The number of complete submissions ranges from 13 (for the \toolname
version of @wwhile@) to 28 (for the \ocaml version of @sumList@), out of
a maximum of 30 per program-tool pair.
%
Collecting more responses per test pair was not possible, as it
would require having students answer the same problem twice (once with
\ocaml and once with \toolname).

\begin{figure}[t]
% \centerline{
% \begin{minipage}{1.2\textwidth}
\includegraphics[width=0.7\linewidth]{user-study-reason.png}
\includegraphics[width=0.7\linewidth]{user-study-fix.png}
% \end{minipage}
% }
% \vspace{3ex}
\caption{A classification of students' explanations and fixes for type
  errors, given either \ocaml's error % message
  or \toolname's
  jump-compressed trace. The students given \toolname's jump-compressed
  trace consistently scored better ($\ge 10\%$) than those given \ocaml's type
  error.}
\label{fig:results-user-study}
\end{figure}

\paragraph{Results}
%
Figure~\ref{fig:results-user-study} summarizes a single annotator's
results, which show that students given \toolname's jump-compressed
trace were consistently more likely to correctly explain
and fix the type error than those given \ocaml's error message.
%
Across each problem the \toolname responses were marked correct
$10-30\%$ more often than the \ocaml responses, which suggests that
the students who had access to \toolname's traces had a better
understanding of the type errors.
%
The measured kappa values were $\kappa = 0.72$ for the explanations and
$\kappa = 0.83$ for the fixes; while there is no formal notion for what
consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
above $0.60$ are often called ``substantial''
agreement~\cite{Landis1977-ey}.
%
%

>>>>

Finally, we used a one-sided Mann-Whitney $U$ test~\cite{Mann1947-fd} to
determine the significance of our results.
%
The null hypothesis was that the responses from students given
\toolname's witnesses were drawn from the same distribution as those
given \ocaml's errors, \ie \toolname had no effect.
%
Since we used a one-sided test, the alternative to the null hypothesis
is that \toolname had a \emph{positive} effect on the responses.
%
We reject the null hypothesis in favor of the alternative if the test
produces a significance level $p < 0.05$, a standard threshold for
determining statistical significance.

@@

\subparagraph{Internal.}
%
We assigned students randomly to two groups. The first was given
\ocaml's errors for @append@ and \hbox{@digitsOfInt@,} and \toolname's trace
for @sumList@ and \hbox{@wwhile@;} the second was given the opposite
assignment of errors and traces. This assignment ensured that: (1) each
student was given \ocaml and \toolname problems; and (2) each student
was given an ``easy'' and ``hard'' problem for both \ocaml and
\toolname. Students without sufficient knowledge of \ocaml could affect
the results, as could the time-constrained nature of an exam. For these
reasons we excluded any answers left blank from our analysis.

@@

\subparagraph{Conclusion.}
%
We collected exams from 60 students, though due to the nature of the
study not every student completed every problem.
%
The number of complete submissions ranges from 13 (for the \toolname
version of @wwhile@) to 28 (for the \ocaml version of @sumList@), out of
a maximum of 30 per program-tool pair.
%
Our results are statistically significant in only 2 out of 8 tests; however,
collecting more responses per test pair was not possible as it
would require having students answer the same problem twice (once with
\ocaml and once with \toolname).

\begin{figure}[t]
% \centerline{
% \begin{minipage}{1.2\textwidth}
\includegraphics[width=0.7\linewidth]{user-study-reason.png}
\includegraphics[width=0.7\linewidth]{user-study-fix.png}
% \end{minipage}
% }
% \vspace{3ex}
\caption{A classification of students' explanations and fixes for type
  errors, given either \ocaml's error message or \toolname's
  jump-compressed trace. The students given \toolname's jump-compressed
  trace consistently scored better ($\ge 10\%$) than those given
  \ocaml's type error. We report the result of a one-sided Mann-Whitney
  $U$ test for statistical significance in parentheses.}
\label{fig:results-user-study}
\end{figure}

\paragraph{Results}
%
The measured kappa values were $\kappa = 0.72$ for the explanations and
$\kappa = 0.83$ for the fixes; while there is no formal notion for what
consititutes strong agreement~\cite{Krippendorff2012-wd}, kappa values
above $0.60$ are often called ``substantial''
agreement~\cite{Landis1977-ey}.
%
Figure~\ref{fig:results-user-study} summarizes a single annotator's
results, which show that students given \toolname's jump-compressed
trace were consistently more likely to correctly explain
and fix the type error than those given \ocaml's error message.
%
Across each problem the \toolname responses were marked correct
$10-30\%$ more often than the \ocaml responses, which suggests that
the students who had access to \toolname's traces had a better
understanding of the type errors;
%
however, only the @append@ tests were statistically significant at
$p < 0.05$.
%
